---
title: "GSS challange"
author: "Angel Ramos Mendiola"
date: "3/15/2021"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(knitr)
GSS <-read.csv(file="GSScsv.csv",header=TRUE)
```
```{r}
GSS$RACE <-as.factor(GSS$RACE)
GSS$BIGBANG <-as.factor(GSS$BIGBANG)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

Is belief in the bigbang theory (BIGBANG) independent of race (RACE)?
If not, then whose view differs from whom??

# Methods

I first made sure that the character variables were turned into factor variables. I then looked into a fisher exact test and a Person exact test to find out what was going on in between the different races.  
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results
```{r}
dd2 <- GSS %>% group_by(RACE,BIGBANG) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=RACE,y=count,fill=BIGBANG))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=RACE,y=prcnt*100,fill=BIGBANG)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

What I see in this plot is that the majority of Black people do not believe in the BIGBANG theory. While in the RACE "Other" the majority does believe in the BIGBANG theory. White people have a 50/50 belief on the theory. Each RACE has a different attitude when it comes to the theory.   
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

### Numerical Descriptive Results
```{r}
table2 <- xtabs(~RACE + BIGBANG, data=GSS)
rowPerc(table2)
colPerc(table2)
```

My observation on the numerical results 80% of the Black population don't believe in the BIGBANG theory. 55% of the Other race does believe in the BIGBANG theory. White people have a 50/50 belief in the BIGBANG theory. There is a big difference in between Black peoples' beliefs in the BIGBANG theory compared to the rest of the groups, while there is not that much of a difference in between the Other and White peoples' beliefs in the BIGBANG theory.  


<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

## Inferential Results
```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
In the inferential results you can see that there is a significance difference in between the P-Value in the Pearson's Chi-squared test which is p-value = 0.01098. While in the Fisher's Exact Test for Count Data test the p-value = 0.00905. The fisher exact test tells me that there is something significant going one, but I still do not know what that is.  

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

# 2X2 test

## A function to run multiple fisher tests
### Upload the fisherXY.R function then source it

```{r}
fisherXY = function(x, y, confL = 0, printAll=FALSE){
  # Inputs: 
  #   x: factor variable, 2 levels
  #   y: factor variable, n levels
  # confL: confidence level, default = 0.95
  #        print only those tables, results if p-value < (1-confL)
  # printAll: overide the 'print only if p-value < (1-confL), and
  #           print all tables and fisher test results
  #
  # Constraint:
  #   x and y must be same length
  #
  # Outputs:
  #   All 2x2 Fisher exact tests
  
  if(confL<0 || confL> 1){
    print('error: confL must be between 0 and 1')
    return(999)
  }
  
  if(!is.factor(x)) x=as.factor(x)
  if(!is.factor(y)) y=as.factor(y)
  
  fisherTable = as.matrix(table(x, y))
  cat('\ncontingency table:\n')
  print(fisherTable)
  cat('\n***All 2X2 tests****\n\n')

  # all 2x2 tests
  levelsY = levels(y)
  levelsX = levels(x)
  
  m = length(levelsX)
  n = length(levelsY)
  
  confLevel = 0.95
  if (confL !=0 ) confLevel=confL
  
  for (i in 1:(m-1))
       for(j in ((i+1):m))
          for (k in 1:(n-1))
             for (l in ((k+1):n)) {
                
                 fisherIndexX = c(i,j)
                 fisherIndexY = c(k,l)
                 
                 fTempTable = as.matrix(fisherTable[fisherIndexX,
                                        fisherIndexY])
                 
                 fisherTest=fisher.test(fTempTable, conf.int = TRUE, 
                              conf.level = confLevel)
                 
                 if((fisherTest$p.value < (1-confLevel) ) || printAll){
                   
                       print(c(levelsX[i], levelsX[j]))
                       print(c(levelsY[k], levelsY[l]))
                       print(fTempTable)
                       print(fisherTest)
                       cat('\n')
                 }
                 
        }
}

```
```{r}
fisherXY(GSS$RACE,GSS$BIGBANG)
```



# Discussion and Conclusion
In conclusion, In the 2x2 test that was ran to see each of the RACES' beliefs on the BIGBANG theory you can see that there  is a significant difference in between Black and Others, there is also a difference in Black and White. 


<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
